---
title: "code for manuscript: Recent lifestyle change impacts sleep and circadian rhythms among the Indigenous peoples of Peninsular Malaysia"
author: "XXX"
date: "2025-07-28"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
rm(list = ls())
library(tidyverse)
library(ggplot2)
library(ggmap)
library(scatterpie) 
library(ggrepel)
library(dplyr)
library(performance)
library(dplyr)
library(corrplot)
library(ggpubr)

library(nationalparkcolors)
pal <- park_palette("Arches",6)

```

## Read in the different datasets
```{r, results = 'hide', fig.show='hide', include=FALSE} 
Medical=read.csv('data/medical.csv')
Traditional=read.csv('data/traditional_lifestyle.csv')
Personal_Information=read.csv('data/personal_information.csv')
Visit_Register <- read.csv("data/visit_register.csv")
Village_Register <- read.csv("data/oa_village_register.csv")
dat1 <- read.csv('data/Final_cleaned_nightsummary_2025-06-12.csv')

# merge
OA_data <- left_join(Medical, Personal_Information, by = "rid", suffix = c(".medical", ".personal"))
OA_data <- left_join(OA_data, Traditional, by = c("rid","vid"))

```

## variable preparation
```{r}

dat2 <- dat1 %>%
  separate(filename, into = c("rid", "letters_ext"), sep = "_") %>%
  separate(letters_ext, into = c("ext1", "ext2"), sep = "\\.")
dat<-merge(OA_data,dat2,by='rid')

# remove under 18
dat$id<-paste(dat$rid,dat$night,sep='_')
dat$age<-as.numeric(difftime(as.Date(dat$med_date), as.Date(dat$date_of_birth), units = "days")) / 365.25 
dat<-subset(dat,age>18)

#convert variables to factors
dat$rid <- factor(dat$rid)
dat$sex_medical <- factor(dat$sex_medical)
dat$house_type <- factor(dat$house_type, levels = c("traditional", "wood", "concrete"))
table(dat$house_type)
dat$housing <- factor(dat$house_type, 
                      levels = c("traditional", "wood", "concrete"), 
                      labels = c("traditional", "nontraditional", "nontraditional"))
table(dat$housing)

dat$electricity_resid <- factor(dat$electricity_resid)
dat$electricity_resid_source___solar <- factor(dat$electricity_resid_source___solar)
dat$electricity_resid_source___generator <- factor(dat$electricity_resid_source___generator)
dat$electricity_resid_source___power_lines <- factor(dat$electricity_resid_source___power_lines)
dat$hh_item___smart_phone <- factor(dat$hh_item___smart_phone)
dat$interview_location_med <- factor(dat$interview_location_med)
dat$wage_past_month <- factor(dat$wage_past_month)

table(dat$electricity_resid_source___solar, useNA = "ifany")
table(dat$electricity_resid_source___generator, useNA = "ifany")
table(dat$electricity_resid_source___power_lines, useNA = "ifany")
table(dat$electricity_resid, useNA = "ifany")

#create new variable for electricity source from 3 variables - solar, generator, powerlines; where more than one occurs, selects higher level
dat$electricity_source <- with(dat, ifelse(
  is.na(electricity_resid_source___power_lines) & 
    is.na(electricity_resid_source___generator) & 
    is.na(electricity_resid_source___solar), NA, # Preserve NAs
  ifelse(
    electricity_resid_source___power_lines == 1, "power_lines",
    ifelse(
      electricity_resid_source___generator == 1, "generator",
      ifelse(
        electricity_resid_source___solar == 1, "solar",
        "none" # Default for no electricity sources
      )
    )
  )
))
# Convert to a factor with appropriate levels
dat$electricity_source <- factor(dat$electricity_source, levels = c("none", "solar", "generator", "power_lines"))
#check that equation worked; NAs for electricity_source should be the same as electricity_resid
```

### FIGURE 1 PANELS
## Get lat/long of village from the village register
```{r}

table(dat$interview_location_med %in% Village_Register$village_id) 
dat$lat <- Village_Register$lat[match(dat$interview_location_med, Village_Register$village_id)]
dat$long <- Village_Register$long[match(dat$interview_location_med, Village_Register$village_id)]

# rename to later variable name used for village code
names(dat)[which(names(dat) == "interview_location_med")] <- "village_id"

# prep the map
medianlat <- median(dat$lat, na.rm = TRUE)
medianlong <- median(dat$long, na.rm = TRUE)

register_google('AIzaSyBqRWaHZTtsvaSOIwDdHjmTVPD-cXvaKrE')
s <- "element:geometry%7Ccolor:0xf5f5f5&style=element:labels%7Cvisibility:off&style=element:labels.icon%7Cvisibility:off&style=element:labels.text.fill%7Ccolor:0x616161&style=element:labels.text.stroke%7Ccolor:0xf5f5f5&style=feature:administrative%7Celement:geometry%7Cvisibility:off&style=feature:administrative.country%7Celement:geometry.stroke%7Ccolor:0x000000%7Cvisibility:on&style=feature:administrative.land_parcel%7Cvisibility:off&style=feature:administrative.land_parcel%7Celement:labels.text.fill%7Ccolor:0xbdbdbd&style=feature:administrative.neighborhood%7Cvisibility:off&style=feature:poi%7Cvisibility:off&style=feature:poi%7Celement:geometry%7Ccolor:0xeeeeee&style=feature:poi%7Celement:labels.text.fill%7Ccolor:0x757575&style=feature:poi.park%7Celement:geometry%7Ccolor:0xe5e5e5&style=feature:poi.park%7Celement:labels.text.fill%7Ccolor:0x9e9e9e&style=feature:road%7Cvisibility:off&style=feature:road%7Celement:geometry%7Ccolor:0xffffff&style=feature:road%7Celement:labels.icon%7Cvisibility:off&style=feature:road.arterial%7Celement:labels.text.fill%7Ccolor:0x757575&style=feature:road.highway%7Celement:geometry%7Ccolor:0xdadada&style=feature:road.highway%7Celement:labels.text.fill%7Ccolor:0x616161&style=feature:road.local%7Celement:labels.text.fill%7Ccolor:0x9e9e9e&style=feature:transit%7Cvisibility:off&style=feature:transit.line%7Celement:geometry%7Ccolor:0xe5e5e5&style=feature:transit.station%7Celement:geometry%7Ccolor:0xeeeeee&style=feature:water%7Celement:geometry%7Ccolor:0xc9c9c9&style=feature:water%7Celement:labels.text.fill%7Ccolor:0x9e9e9e&size=480x360"
c(medianlong, medianlat)

map <- get_googlemap(center=c(102,4), zoom = 7, scale = 1, style = s)
m <- ggmap(map)
```

## Map just showing the sampling locations
```{r}
sample_local_plot<- m +
 geom_point(data = dat, aes(x = long, y = lat), fill='steelblue', color="black", alpha = 0.2, size=3, shape=21) +
  xlab("Longitude") +
  ylab("Latitude") +
  xlim(100.5, 103.5) +
  ylim(2, 6.5) +
  theme_bw(13); sample_local_plot
```

## Wage labor participation, non-trad housing and powerline electricity in each village (Figure 1A)
```{r} 

dat$electricity_source2<-ifelse(dat$electricity_source=='power_lines',1,0)
dat$house_type2<-ifelse(dat$house_type=='traditional',0,1)

unique_ids<-unique(dat[,c('wage_past_month','electricity_source2','house_type2','village_id','tid')])

OA_wage_by_village<- unique_ids %>%
  group_by(village_id) %>%
  summarize(
    has_wage = sum(wage_past_month == 1, na.rm = TRUE),  # Count people who said yes
    asked_wage = sum(!is.na(wage_past_month)),           # Count people who were asked 
    percent_with_wage = has_wage / asked_wage * 100,
    no_wage= asked_wage-has_wage
  )

OA_house_by_village<- unique_ids %>%
  group_by(village_id) %>%
  summarize(
    has_trad = sum(house_type2 == 1, na.rm = TRUE),  # Count people who said yes
    asked_house = sum(!is.na(house_type2)),           # Count people who were asked 
    percent_with_trad = has_trad / asked_house * 100,
    no_trad= asked_house-has_trad
  )


OA_elec_by_village<- unique_ids %>%
  group_by(village_id) %>%
  summarize(
    has_elec = sum(electricity_source2 == 1, na.rm = TRUE),  # Count people who said yes
    asked_elec = sum(!is.na(electricity_source2)),           # Count people who were asked 
    percent_with_elec = has_elec / asked_elec * 100,
    no_elec= asked_elec-has_elec
  )
    
# Create sample data
OA_village_local <- dat %>%
  group_by(village_id) %>%
  summarise(
    lat = mean(lat, na.rm = TRUE),
    long = mean(long, na.rm = TRUE))

# prep to plot wage labor
OA_village_plot<-merge(OA_wage_by_village, OA_village_local, by="village_id") %>%
  na.omit()
OA_village_plot$size<- OA_village_plot$asked_wage/max(OA_village_plot$asked_wage)
OA_village_plot<-subset(OA_village_plot,asked_wage>4)

# shaded version of wage labor
wage_plot2<- m +
 geom_point(data = OA_village_plot, aes(x = long, y = lat, fill=percent_with_wage,size=asked_wage), color="black", alpha = 0.5, shape=21) +
  xlab("Longitude") +
  ylab("Latitude") +
  xlim(100.5, 103.5) +
  ylim(2, 6.5) +
  theme_bw(13)+scale_fill_gradient(low="blue", high="red",limits = c(0, 100))+labs(size='N',fill = "% wage labor"); wage_plot2

# prep to plot electricity
OA_village_plot<-merge(OA_elec_by_village, OA_village_local, by="village_id") %>%
  na.omit()
OA_village_plot$size<- OA_village_plot$asked_elec/max(OA_village_plot$asked_elec)
OA_village_plot<-subset(OA_village_plot,asked_elec>4)

# shaded version of electricity
elec_plot2<- m +
 geom_point(data = OA_village_plot, aes(x = long, y = lat, fill=percent_with_elec,size=asked_elec), color="black", alpha = 0.5, shape=21) +
  xlab("Longitude") +
  ylab("Latitude") +
  xlim(100.5, 103.5) +
  ylim(2, 6.5) +
  theme_bw(13)+scale_fill_gradient(low="blue", high="red",limits = c(0, 100))+labs(size='N',fill = "% power lines"); elec_plot2

# prep to plot house type
OA_village_plot<-merge(OA_house_by_village, OA_village_local, by="village_id") %>%
  na.omit()
OA_village_plot$size<- OA_village_plot$asked_house/max(OA_village_plot$asked_house)
OA_village_plot<-subset(OA_village_plot,asked_house>4)

# shaded version of house type
house_plot2<- m +
 geom_point(data = OA_village_plot, aes(x = long, y = lat, fill=percent_with_trad,size=asked_house), color="black", alpha = 0.5, shape=21) +
  xlab("Longitude") +
  ylab("Latitude") +
  xlim(100.5, 103.5) +
  ylim(2, 6.5) +
  theme_bw(13)+scale_fill_gradient(low="blue", high="red",limits = c(0, 100))+labs(size='N',fill = "% non traditional\n house"); house_plot2

```
## Prepare environment for local climate and sunrise/set
```{r}
library(suncalc)
library(ecmwfr)
library(ncdf4)
```

## Extract and calculate local sunrise/sunset (Figure 1D)
```{r}
lat <- medianlat
lon <- medianlong
# Define years and location
years <- 2020:2024
# Create date sequence
years <- 2020:2024  # Jan 2020 – Dec 2024
dates <- bind_rows(lapply(years, function(y) {
  tibble(
    date = seq.Date(as.Date(paste0(y, "-01-01")), as.Date(paste0(y, "-12-31")), by="day"),
    year = y
  )
}))
# Get daily sunrise/sunset times
sun_times <- getSunlightTimes(
  date = dates$date, lat = lat, lon = lon, keep = c("sunrise", "sunset"), tz = "Asia/Kuala_Lumpur"
) %>%
  left_join(dates, by = "date") %>%
  mutate(
    doy = yday(date),
    month = month(date, label=TRUE, abbr=TRUE),
    sunrise_hour = hour(sunrise) + minute(sunrise)/60,
    sunset_hour  = hour(sunset) + minute(sunset)/60
  )
#calculate averages
monthly_sun_stats <- sun_times %>%
  group_by(month) %>%
  summarise(
    mean_sunrise = mean(sunrise_hour, na.rm=TRUE),
    se_sunrise   = sd(sunrise_hour, na.rm=TRUE) / sqrt(n()),
    mean_sunset  = mean(sunset_hour, na.rm=TRUE),
    se_sunset    = sd(sunset_hour, na.rm=TRUE) / sqrt(n()),
    .groups='drop'
  ) %>%
  mutate(month = factor(month, levels = month.abb, ordered = TRUE))
#plot sunrise/sunset
ggplot(monthly_sun_stats, aes(x = month)) +
  # Confidence ribbons
  geom_ribbon(aes(
    ymin = mean_sunrise - se_sunrise,
    ymax = mean_sunrise + se_sunrise,
    fill = "Sunrise"
  ), alpha = 0.3) +
  geom_ribbon(aes(
    ymin = mean_sunset - se_sunset,
    ymax = mean_sunset + se_sunset,
    fill = "Sunset"
  ), alpha = 0.3) +

  # Mean lines
  geom_line(aes(y = mean_sunrise, color = "Sunrise", group = 1), size = 1.0) +
  geom_line(aes(y = mean_sunset, color = "Sunset", group = 1), size = 1.0) +

  # Axes
  scale_y_continuous(
    limits = c(6, 20),
    breaks = seq(6, 20, by = 6),
    name = "Hour of day"
  ) +
  scale_x_discrete(expand = c(0, 0)) +  # remove padding between months
  scale_color_manual(values = c("Sunrise" = "goldenrod2", "Sunset" = "steelblue")) +
  scale_fill_manual(values = c("Sunrise" = "gold", "Sunset" = "lightblue")) +

  # Labels
  labs(
    title = "Monthly Sunrise and Sunset Times (2020–2024)",
    x = "Month",
    color = "",
    fill = ""
  ) +

  # Legend
  guides(
    color = guide_legend(order = 2),
    fill  = guide_legend(order = 2)
  ) +

  # Matching theme
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.x = element_line(color = "grey85"),
    panel.grid.minor.x = element_line(color = "grey90"),  # adds those subtle lines between months
    panel.grid.major.y = element_line(color = "grey85"),
    panel.grid.minor.y = element_line(color = "grey90"),
    axis.text = element_text(family = "sans"),             # match fonts
    axis.title = element_text(family = "sans"),
    plot.title = element_text(hjust = 0.5, family = "sans"),
    plot.margin = margin(t = 10, r = 15, b = 10, l = 15)    # clean margins
  )
```

## Local climate plot (Figure 1E)
```{r}
climate <- read.csv('data/Malaysia_monthly_climate_2020-2024.csv')
#convert Kelvins to Celsius
climate_clean <- climate %>%
  mutate(
    date = ymd_hms(valid_time),
    year = year(date),
    month_num = month(date),
    month_label = month(date, label = TRUE, abbr = TRUE),
    temp_C = t2m - 273.15
  ) %>%
  filter(year >= 2020, year <= 2024)
#mean ± SE for each month across all hourly datapoints
climate_summary_raw <- climate_clean %>%
  group_by(month_num, month_label) %>%
  summarise(
    mean_temp = mean(temp_C, na.rm = TRUE),
    se_temp = sd(temp_C, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )
print(climate_summary_raw)
summary(climate_summary_raw$se_temp)

#Plot monthly averages and SE loess lines
ggplot(climate_summary_raw, aes(x = month_num, y = mean_temp)) +
  geom_ribbon(aes(ymin = mean_temp - se_temp, ymax = mean_temp + se_temp),
              fill = "#56B4E9", alpha = 0.5) +
  geom_line(color = "black", size = 1.2) +
  scale_x_continuous(
    breaks = 1:12,
    labels = levels(climate_summary_raw$month_label)
  ) +
  labs(
    title = "Monthly Mean Temperature (2020–2024)",
    x = "Month",
    y = "Temperature (°C)"
  ) +
  theme_minimal(base_size = 14)

```

### FIGURE 2 PANELS
## Variability; Age vs sleep outcomes (Figure 2c/2d)
```{r}
# awakenings = number_of_awakenings
# offset = wakeup
# duration = SleepDurationInSpt
# onset = sleeponset
# SRI = SleepRegularityIndex
# WASO = WASO

dat2<-dat[,c('rid','number_of_awakenings','wakeup','SleepDurationInSpt','sleeponset','SleepRegularityIndex1','WASO','age','id')]
apply(dat[,2:7],2,function(x) mean(x,na.rm=T))
apply(dat[,2:7],2,function(x) sd(x,na.rm=T))


# estimate variability
out<-matrix(nrow=7,ncol=5)
for (i in 2:7){
model <- aov(dat2[,i] ~ rid, data = dat2)
anova_table <- summary(model)[[1]]
out[i,1] <- anova_table[1, "Sum Sq"]
out[i,2] <- anova_table[2, "Sum Sq"]
SS_total <- out[i,1] + out[i,2]
out[i,3] <- out[i,1] / SS_total
out[i,4]  <- out[i,2] / SS_total
out[i,5]  <- unlist(anova_table[4])[1]
}

# standardize
dat2$Awakenings<-scale(dat2$number_of_awakenings)
dat2$Offset<-scale(dat2$wakeup)
dat2$Duration<-scale(dat2$SleepDurationInSpt)
dat2$Onset<-scale(dat2$sleeponset)
dat2$SRI<-scale(dat2$SleepRegularityIndex1)
dat2$WASO<-scale(dat2$WASO)

# plot an example
offset_summary <- dat2 %>%
  group_by(rid, age) %>%
  summarize(
    mean_offset = mean(wakeup, na.rm = TRUE),
    n_obs = n(),
    .groups = "drop"
  )

time_labels <- function(x) {
  hour <- ifelse(x >= 24, x - 24, x)
  h <- floor(hour)
  m <- round((hour - h) * 60)
  sprintf("%02d:%02d", h, m)
}

# Example ggplot with effect of age on sleep offset
ggplot(offset_summary, aes(x = age, y = mean_offset)) +
  geom_point(alpha=0.3,color=pal[1]) +
  scale_y_continuous(
    name = "Offset",
    labels = time_labels
  ) +
  labs(x = "Age") +
  theme_minimal(13)+stat_smooth(method='lm',se=FALSE,size=2,color=pal[1])

## fig 2C - effect of age on sleep outcomes
dat3 <- dat2[,c('Awakenings','Offset','Duration','Onset','SRI','WASO','age','id')] %>%
  pivot_longer(!id & !age, names_to = "variable", values_to = "value")

dat3$variable2<-fct_relevel(as.factor(dat3$variable), 
            "Offset", "Onset", "SRI", 
            "Awakenings", "WASO", "Duration")

ggplot(dat3,aes(x=age, y=value,color=variable2)) +
    stat_smooth(method='lm',se=FALSE) +
    theme_minimal(13) +
    xlab("Age") +
    ylab("Standardized value") +
    scale_color_manual(values=pal)+guides(fill = guide_legend(title = "Outcome"))


```

## Corrplot (Figure 2B)
```{r}
# awakenings = number_of_awakenings
# offset = wakeup
# duration = SleepDurationInSpt
# onset = sleeponset
# SRI = SleepRegularityIndex
# WASO = WASO

dat2<-dat[,c('number_of_awakenings','wakeup','SleepDurationInSpt','sleeponset','SleepRegularityIndex1','WASO')]
names(dat2)<-c('Awakenings','Offset','Duration','Onset','SRI','WASO')
corr_matrix<-cor(dat2,use='pairwise.complete.obs')

cor.mtest <- function(mat, conf.level = 0.95) {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], conf.level = conf.level)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  return(p.mat)
}

# Compute p-values
p.mat <- cor.mtest(dat2)

# corrplot(corr_matrix, method='ellipse',type = 'lower',  tl.col = 'black',
#         cl.ratio = 0.2, tl.srt = 45, col = COL2('PuOr', 10),diag=FALSE)
#fig 2b
corrplot::corrplot(corr_matrix, method='ellipse',type = 'lower',  tl.col = 'black',
         cl.ratio = 0.2, tl.srt = 45, col = colorRampPalette(c("white", pal[3],pal[5]))(10),diag=FALSE)
```

## Density (Figure 2E)
```{r}
# awakenings = number_of_awakenings
# offset = wakeup
# duration = SleepDurationInSpt
# onset = sleeponset
# SRI = SleepRegularityIndex
# WASO = WASO

dat2<-dat[,c('number_of_awakenings','wakeup','SleepDurationInSpt','sleeponset','SleepRegularityIndex1','WASO','id','sex_medical')]
names(dat2)[1:6]<-c('Awakenings','Offset','Duration','Onset','SRI','WASO')

dat3 <- dat2 %>%
  pivot_longer(!id & !sex_medical, names_to = "variable", values_to = "value")

library(ggridges)

dat3$variable2<-fct_relevel(as.factor(dat3$variable), 
            "Offset", "Onset", "SRI", 
            "Awakenings", "WASO", "Duration")

ggplot(dat3,aes(x=value, fill=variable2,y=sex_medical)) +
    geom_density_ridges() +
   theme_ridges() + 
    theme(legend.position = "none") +
    theme_minimal(13) +
    xlab("Value") +
    ylab("Density") +
    facet_wrap(~variable2,scales='free')+scale_fill_manual(values=pal)+theme(legend.position='none')
```

## Prepare environment for multilevel models
```{r}
library(lme4)
library(broom.mixed)
library(gridExtra)
library(nationalparkcolors)
library(egg)
library(merTools)
library(emmeans)
```

## Prepare Multilevel models
```{r}
lmm_WASO <- lmer(scale(log(WASO + 1)) ~ wage_past_month + housing + electricity_source + hh_item___smart_phone + scale(age) + sex + (1|rid), data = dat)

glmm_awakenings <- glmer(
  number_of_awakenings ~ wage_past_month + housing + electricity_source + hh_item___smart_phone + scale(age) + sex + (1|rid),
  family = poisson(link = "log"), 
  data = dat,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

lmm_awakenings <- lmer(
  scale(number_of_awakenings) ~ wage_past_month + housing + electricity_source + hh_item___smart_phone + scale(age) + sex + (1|rid),
  data = dat,
)

lmm_duration <- lmer(scale(SleepDurationInSpt) ~ wage_past_month + housing * electricity_source + hh_item___smart_phone + scale(age) + sex + (1|rid), data = dat)

ggplot(dat,aes(x=age, y=wakeup))+geom_point()+ stat_smooth(method="lm")
lmm_wake <- lmer(scale(wakeup) ~ wage_past_month  + housing + electricity_source + hh_item___smart_phone + scale(age) + sex + (1|rid), data = dat)

lmm_onset <- lmer(scale(sleeponset) ~ wage_past_month + housing + electricity_source + hh_item___smart_phone + scale(age) + sex + (1|rid), data = dat)

lmm_regularity <- lmer(scale(SleepRegularityIndex1) ~ wage_past_month + housing + electricity_source + hh_item___smart_phone + scale(age) + sex + (1|rid), data = dat)
```

## Extract fixed effects from lme4 models
```{r}
# Function to extract fixed effects from lme4 models
extract_fixed_effects <- function(model, model_name, category) {
  # Extract fixed effects and their standard errors
  fixed_effects <- tidy(model, effects = "fixed", conf.int = TRUE)
  
  # Format the data
  fixed_effects_df <- fixed_effects %>%
    select(term, estimate, std.error, conf.low, conf.high) %>%
    mutate(
      model = model_name,
      category = category,
      # Rename the intercept term to make it clearer in the plot
      variable = ifelse(term == "(Intercept)", "Intercept", term)
    ) %>%
    # If you want to exclude the intercept, uncomment the next line
    # filter(term != "(Intercept)") %>%
    select(model, variable, estimate, std.error, conf.low, conf.high, category)
  
  return(fixed_effects_df)
}
# Function to extract the dependent variable name from a model formula
get_dependent_var <- function(model) {
  # Extract the formula
  form <- formula(model)
  # Get the response variable (LHS of the formula)
  response_var <- all.vars(form)[1]
  return(response_var)
}

# Function to extract fixed effects from lme4 models
extract_fixed_effects <- function(model, model_name) {
  # Extract dependent variable name for labeling
  dep_var <- get_dependent_var(model)
  
  # Extract fixed effects and their standard errors
  fixed_effects <- tidy(model, effects = "fixed", conf.int = TRUE)
  
  # Format the data
  fixed_effects_df <- fixed_effects %>%
    dplyr::select(term, estimate, std.error, conf.low, conf.high) %>%
    mutate(
      model = model_name,
      dependent_var = dep_var,
      # Rename the intercept term to make it clearer in the plot
      variable = ifelse(term == "(Intercept)", "Intercept", term)
    ) %>%
    # If you want to exclude the intercept, uncomment the next line
    # filter(term != "(Intercept)") %>%
    dplyr::select(model, variable, estimate, std.error, conf.low, conf.high, dependent_var)
  
  return(fixed_effects_df)
}
all_model_results <- bind_rows(list(
  extract_fixed_effects(lmm_onset, "Onset"),
  extract_fixed_effects(lmm_wake, "Offset"),
  extract_fixed_effects(lmm_regularity, "SRI"),
  extract_fixed_effects(lmm_WASO, "WASO"),
  extract_fixed_effects(lmm_awakenings, "Awakenings"),
  extract_fixed_effects(lmm_duration, "Duration")
))
```

## Create function to generate forest plots
```{r}
create_forest_plot <- function(data, title, col_choice) {
  ggplot(data, aes(x = estimate, y = variable, color = model)) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
    geom_point(aes(alpha=sig), position = position_dodge(width = 0.5), size = 3) +
    geom_errorbarh(aes(xmin = conf.low, xmax = conf.high, alpha=sig), 
                   position = position_dodge(width = 0.5),
                   height = 0.2, linewidth = 0.5) +
    scale_color_manual(values = col_choice) +
    scale_alpha(guide = 'none', range = c(0.5,1))+
    labs(
      title = title,
      x = "Coefficient",
      y = NULL,
      color = ""
    ) +
    theme_bw() +
    theme(
      legend.position = "bottom",
      panel.grid.minor = element_blank(),
      plot.title = element_text(hjust = 0.5, face = "bold"),
      axis.text.y = element_text(hjust = 0)
    )
}

all_model_results <- all_model_results[-which(all_model_results$variable == "Intercept"),]
all_model_results$variable[which(all_model_results$variable == "scale(age)")] <- "Age"
all_model_results$variable[which(all_model_results$variable == "electricity_sourcegenerator")] <- "Electricity (generator)"
all_model_results$variable[which(all_model_results$variable == "electricity_sourcepower_lines")] <- "Electricity (powerlines)"
all_model_results$variable[which(all_model_results$variable == "electricity_sourcesolar")] <- "Electricity (solar)"
all_model_results$variable[which(all_model_results$variable == "hh_item___smart_phone1")] <- "Has phone"
all_model_results$variable[which(all_model_results$variable == "housingnontraditional")] <- "Housing (non-traditional)"
all_model_results$variable[which(all_model_results$variable == "sex1")] <- "Sex"
all_model_results$variable[which(all_model_results$variable == "wage_past_month1")] <- "Participates in wage labor"
all_model_results$sig <- ifelse(all_model_results$conf.low < 0 & all_model_results$conf.high < 0 |
                                  all_model_results$conf.low > 0 & all_model_results$conf.high > 0,
                                1, .5)


all_model_results$variable <- factor(all_model_results$variable, 
                                     levels=rev(c("Age", "Sex", "Participates in wage labor", "Has phone", "Housing (non-traditional)", "Electricity (solar)", "Electricity (generator)", "Electricity (powerlines)")))
```

### FIGURE 3
## forest plots
```{r}
# Create three forest plots based on model groups
plot1 <- create_forest_plot(
  filter(all_model_results, model %in% c("Onset", "Offset", "SRI")),
  "Timing", col_choice = pal[1:3]
)

plot2 <- create_forest_plot(
  filter(all_model_results, model %in% c("WASO", "Awakenings")),
  "Quality", col_choice = pal[4:5]
)
plot2 <- plot2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank())


plot3 <- create_forest_plot(
  filter(all_model_results, model %in% c("Duration")),
  "Quantity", col_choice = pal[6]
)
plot3 <- plot3 + theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank())

combined_plot <- egg::ggarrange(plots=list(plot1, plot2, plot3),ncol=3)
```

